Ground state energy of quantum dots using the coupled cluster method

### 2018-06-03 project_2/article/article.tex
Begun intermediates control. Fock operator correction.

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
            \label{eq:doubles_amplitude_1}
        \end{align}
        to find an expression that can be used to solve for $t_{ij}^{ab}$.
        By employing the Baker-Campbell-Haussdorf (BCH) expansion,
        while setting $T= T_2$, we find
        \begin{equation}
        \begin{aligned}
        		\bar{H} 
        			=& \Big(
        					H_N + H_NT_2 
        					+ \frac{1}{2}H_NT_2^2 \\
        				  	&\ + \frac{1}{6}H_NT_2^3 
        				   + \frac{1}{24} H_NT_2^4
        				 \Big)_C.
        \end{aligned}
        \label{eq:normal_order_expansion}
        \end{equation}
		The subscript $c$ indicates that only those terms in which the 
		Hamiltonian is connected\footnote{In a Wick's theorem sense}
		to every cluster operator on its right should be included. Since
		the Hamiltonian contains at most four annihilation and creation 
		operators, $H_N$ can connect to at most four cluster operators
		at once. Therefor, the BCH expansion must truncate at the
		fourth-order terms.

		Now comes the rather tedious task of evaluating all the terms 
		that arises from inserting \autoref{eq:normal_order_expansion}
		into \autoref{eq:doubles_amplitude_1}. This can be done by applying
		Wick's generalised theorem, but the task is a daunting and streneous one.
		A few example computations of how this can be done is included in
		\autoref{app:wickOnAmplitude}.
	 	Instead of doing it in this manner, we employ the second quantisation library
	 	from SymPy instead\footnote{This is also more in the spirit of this project,
	 	as it is within the realm of \emph{Computational} Physics.}.
		The CCD amplitude equation, from SymPy computation, is 
		\begin{equation}
		\begin{aligned}
			0 =& u^{ab}_{ij} + f_{bc}t^{ac}_{ij}P(ab) - f_{kj}t^{ab}_{ik}P(ij) \\
				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
				 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}. 
		\label{eq:CCD_amp_1}
		\end{aligned}
		\end{equation}

		\subsubsection{The Iterative Scheme}
		Now we pick only diagonal elements of $f$ to be part of the unperturbed 
		Hamiltonian and consider the rest of the terms a perturbation. 
		The second and third terms in \autoref{eq:CCD_amp_1} can now be 
		rewritten,
		\begin{equation}
		\begin{aligned}
			f_{bc}&t_{ij}^{ab}P(ab) - f_{kj}t_{ik}^{ab}P(ab) \\
				&= f_{bb}t_{ij}^{ab} - f_{aa}t_{ij}^{ba} - f_{jj}t_{ij}^{ab} + f_{ii}t_{ji}^{ab} \\
				&= (f_{aa} + f_{bb} - f_{ii} - f_{jj})t_{ij}^{ab} \\
				&= (\epsilon_a + \epsilon_b - \epsilon_i - \epsilon_j)t_{ij}^{ab} \\
				&= -(\epsilon_i + \epsilon_j - \epsilon_a - \epsilon_b)t_{ij}^{ab} \\				
				&= -D_{ij}^{ab}t_{ij}^{ab}.
		\end{aligned}
		\end{equation}

		Now, we define a new function consisting of all of \autoref{eq:CCD_amp_1} 
		except term number two and three,
		\begin{equation}
		\begin{aligned}		
			g(u, \tau) =&
				    \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
				 & \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}, 
		\end{aligned}
		\end{equation}

		so that $D_{ij}^{ab}t_{ij}^{ab} = g(u, \tau)$. This allows us to define an iterative scheme,
		\begin{equation}
			\tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
		\end{equation}
		with the initial guess
		\begin{equation}
			\tau^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
		\end{equation}

		\subsubsection{Intermediate Computations}
		Looking closely at the amplitude equation in (\ref{eq:CCD_amp_1}) one might come
		to realize that it contains that this equation contains many of the same structures in
		several of the terms. This warrants the search for an algebraic transformation of the
		CCD amplitude equation that has the potential to reduce the amount of floating point
		operations needed to compute it. As it turns out, such terms exist and they will decrease
		the computing time necessary by an order of magnitude. We will define the following
		"intermediates",
		\begin{align}
			 \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
    			 \chi^b_c &= u^{bm}_{cm} - \frac{1}{2} t^{bd}_{mn} u^{mn}_{cd} \\
   			 \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} \\
    			 \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd}.
		\end{align}
		These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
		\begin{equation}
		\begin{aligned}
			0 =& \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij} - t^{ab}_{im}u^{mn}_{jn}P(ij)
      		+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ac}_{ij}\chi^b_c P(ab) \\
      	  &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
      		+ t^{ab}_{in}\chi^n_jP(ij) + u^{ab}_{ij}.
		\end{aligned}
		\end{equation}
		The importance of this "trick" will become apparent in due time.

>>>>

    \subsection{Coupled cluster amplitude equations}
        In order for us to solve the energy equation using the coupled cluster
        approximation we need to figure out what the cluster amplitudes,
        $t_{ij\dots}^{ab\dots}$, are. This is done by projecting
        \autoref{eq:clean_schrodinger} onto an excited Slater determinant, i.e.,
        \begin{align}
            \bra{\Phi_{ij\dots}^{ab\dots}}e^{-T}He^{T}\kslat
            = 0.
        \end{align}
        Note that in the amplitude equations we can use both the regular and the
        normal ordered Hamiltonian. They are equal as the reference energy term
        disappears due to \autoref{eq:excited_overlap}. The order of the
        excitation in the projection determines the order of the amplitudes you
        will find. In our case we are only interested in the second order
        ampltiudes found in the doubles approximation, hence we will solve the
        equation
        \begin{align}
            \bra{\Phi_{ij}^{ab}}e^{-T}H_N e^{T}\kslat = 0,
            \label{eq:doubles_amplitude_1}
        \end{align}
        to find an expression that can be used to solve for $t_{ij}^{ab}$.
        By employing the Baker-Campbell-Haussdorf (BCH) expansion,
        while setting $T= T_2$, we find
        \begin{equation}
        		\bar{H} 
        			= \Big(
        					H_N + H_NT_2 
        					+ \frac{1}{2}H_NT_2^2 \Big)_C.
        \label{eq:normal_order_expansion}
        \end{equation}
		The subscript $c$ indicates that only those terms in which the 
		Hamiltonian is connected\footnote{In a Wick's theorem sense}
		to every cluster operator on its right should be included. Since
		the Hamiltonian contains at most four annihilation and creation 
		operators, $H_N$ can connect to at most four cluster operators
		at once. Therefor, the BCH expansion must truncate at the
		fourth-order terms.

		Now comes the rather tedious task of evaluating all the terms 
		that arises from inserting \autoref{eq:normal_order_expansion}
		into \autoref{eq:doubles_amplitude_1}. This can be done by applying
		Wick's generalised theorem, but the task is a daunting and streneous one.
		A few example computations of how this can be done is included in
		\autoref{app:wickOnAmplitude}.
	 	Instead of doing it in this manner, we employ the second quantisation library
	 	from SymPy instead\footnote{This is also more in the spirit of this project,
	 	as it is within the realm of \emph{Computational} Physics.}.
		The CCD amplitude equation, from SymPy computation, is 
		\begin{equation}
		\begin{aligned}
			0 =& u^{ab}_{ij} + f^b_c t^{ac}_{ij}P(ab) - f^k_jt^{ab}_{ik}P(ij) \\
				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
				 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}. 
		\label{eq:CCD_amp_1}
		\end{aligned}
		\end{equation}

		\subsubsection{The Iterative Scheme}
		At first, we pick only diagonal elements of $f$ to be part of the unperturbed 
		Hamiltonian and consider the rest of the terms a perturbation. 
		The second and third terms in \autoref{eq:CCD_amp_1} can now be 
		rewritten,
		\begin{equation}
		\begin{aligned}
			f^b_c&t_{ij}^{ab}P(ab) - f^k_j t_{ik}^{ab}P(ij) \\
				&\to f^b_bt_{ij}^{ab} - f^a_at_{ij}^{ba} - f^j_jt_{ij}^{ab} + f^i_it_{ji}^{ab} \\
				&= (f^a_a + f^b_b - f^i_i - f^j_j)t_{ij}^{ab} \\
				&= (\epsilon_a + \epsilon_b - \epsilon_i - \epsilon_j)t_{ij}^{ab} \\
				&= -(\epsilon_i + \epsilon_j - \epsilon_a - \epsilon_b)t_{ij}^{ab} \\				
				&= -D_{ij}^{ab}t_{ij}^{ab}.
		\end{aligned}
		\end{equation}

		The we can define a new function consisting of all of \autoref{eq:CCD_amp_1} 
		except term number two and three,
		\begin{equation}
		\begin{aligned}		
			g(u, \tau) =&
					u^{ab}_{ij} + \tilde{f}^b_c t^{ac}_{ij}P(ab) - \tilde{f}^k_jt^{ab}_{ik}P(ij) \\
				 +& \frac{1}{4}t^{cd}_{ij} t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}t^{cd}_{ij} u^{ab}_{cd} \\
				 +& \frac{1}{2}t^{cd}_{jm} t^{ab}_{in} u^{mn}_{cd} P(ij) - \frac{1}{2}t^{ac}_{nm} t^{bd}_{ij} u^{nm}_{cd} P(ab) \\
				 +& t^{ac}_{im} t^{bd}_{jn} u^{mn}_{cd} P(ij) + t^{ac}_{im} u^{bm}_{jc} P(ab) P(ij) \\
				 +& \frac{1}{2}t^{ab}_{mn} u^{mn}_{ij}, 
		\end{aligned}
		\end{equation}

		where $\tilde{t}$ are the non-diagonal parts of the Fock matrix.

		Now we have $D_{ij}^{ab}t_{ij}^{ab} = g(u, \tau)$. 
		This allows us to define an iterative scheme,
		\begin{equation}
			\tau^{(k + 1)} = \frac{g(u, \tau^{(k)})}{D_{ij}^{ab}},
		\end{equation}
		with the initial guess
		\begin{equation}
			\tau^{(0)} = \frac{u_{ij}^{ab}}{D_{ij}^{ab}}.
		\end{equation}

		\subsubsection{Intermediate Computations}
		Looking closely at the amplitude equation in (\ref{eq:CCD_amp_1}) one might come
		to realize that it contains that this equation contains many of the same structures in
		several of the terms. This warrants the search for an algebraic transformation of the
		CCD amplitude equation that has the potential to reduce the amount of floating point
		operations needed to compute it. As it turns out, such terms exist and they will decrease
		the computing time necessary by an order of magnitude. We will define the following
		"intermediates",
		\begin{align}
			 \chi^{ab}_{cd} &= \frac{1}{4}t^{ab}_{mn} u^{mn}_{cd} + \frac{1}{2}u^{ab}_{cd} \\
    			 \chi^a_d &= \frac{1}{2} t^{ac}_{nm} u^{nm}_{cd} \\
   			 \chi^{bm}_{jc} &= u^{bm}_{jc} + \frac{1}{2}t^{bd}_{jn}u^{mn}_{cd} \\
    			 \chi^n_j &= \frac{1}{2}t^{cd}_{jm} u^{mn}_{cd}.
		\end{align}
		These intermediate structures will allow us to rewrite \autoref{eq:CCD_amp_1} to,
		THIS REMAINS TO BE CHECKED!! 
		\begin{equation}
		\begin{aligned}
			0 =& \frac{1}{2}t^{ab}_{mn}u^{mn}_{ij} - t^{ab}_{im}u^{mn}_{jn}P(ij)
      		+ t^{cd}_{ij}\chi^{ab}_{cd} + t^{ac}_{ij}\chi^b_c P(ab) \\
      	  &+ t^{ac}_{im}\chi^{bm}_{jc}P(ab)P(ij)
      		+ t^{ab}_{in}\chi^n_jP(ij) + u^{ab}_{ij}.
		\end{aligned}
		\end{equation}
		The importance of this "trick" will become apparent in due time.
